Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-09-14 article.tex
Added cantilever beam geometry image

Finite volume formulations for solid mechanics are heavily influenced by their fluid mechanics counterparts, favouring segregated implicit and fully explicit methods.
Segregated approaches, where the governing momentum equation is temporarily decomposed into scalar component equations, offer memory efficiency and simplicity of implementation, but the outer coupling Picard iterations often suffer from slow convergence.
Explicit formulations are straightforward to implement and offer superior robustness but are only efficient for high-speed dynamics, where the physics requires small time increments.
In contrast, the finite element community commonly employs Newton-Raphson-type solution algorithms, which necessitate repeated assembly of the Jacobian matrix and solution of the resulting block-coupled non-diagonally dominant linear system.
A disadvantage of traditional Newton-based approaches is that they typically require explicit formation and storage of the Jacobian matrix, which can be computationally expensive and memory-intensive.
A further disadvantage from a finite volume perspective is that extending existing code frameworks from segregated algorithms to a coupled Newton-Raphson-type approach is challenging in terms of the required assembly, storage, and solution of the resulting block-coupled system.
In addition, the derivation of the true Jacobian matrix is non-trivial.
Consequently, similar block-coupled solution finite volume methods are rare in the literature \citep{Das2012, Cardiff2016, Castrillo2024}.
The motivation of the current work is to seek (or exceed) the robustness and efficiency of block-coupled Newton-Raphons approaches in a way that can be easily incorporated into existing segregated solution frameworks.
To this end, the current article examines the efficacy of \emph{Jacobian-free} Newton-Krylov methods, where the quadratic convergence of Newton methods can potentially be achieved without deriving, assembling and storing the exact Jacobian.

@@

A crucial aspect of ensuring the efficiency and robustness of the Jacobian-free Newton-Krylov method is the choice of a suitable preconditioner for the Krylov iterations.
This preconditioner is often derived from the exact Jacobian matrix in traditional Newton methods.
However, the Jacobian-free approach does not allow direct access to the full Jacobian matrix, necessitating an alternative strategy to approximate its action.
To this end, and to extend existing segregated frameworks, we propose using a compact-stencil approximate Jacobian as the preconditioner. This approximate Jacobian corresponds to the matrix typically employed in segregated approaches; similar approaches are successful in fluid mechanics applications \citep{NishikawaPaper, nonNewtonianJFNKPaper}; however, it is unclear if such an approach is suitable for solid mechanics - a question which we hope to answer in this work.
By leveraging this compact-stencil approximate Jacobian, we aim to effectively precondition the Krylov iterations, enhancing convergence while maintaining the memory and computational savings that define the Jacobian-free and segregated methods.
Similarly, if such an approach is efficient, it would naturally fit into existing segregated frameworks, as existing matrix storage and assembly can be reused.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Cell-Centred Finite Volume Discretisation}
\label{sec:discretisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this work, a nominally second-order cell-centred finite volume discretisation is employed, as described previously, for example, \citep{cardiff_lagrangian_2017, Batistic, Tukovic, Jasak}.
Consequently, only a summary of the discretisation is presented below.

@@

The conservation equation (Equations \ref{eqn:eqn:momentum_lingeom}, \ref{eqn:eqn:momentum_TL}, or \ref{eqn:eqn:momentum_UL}) is applied to each cell (control volume) in the computational mesh and discretised in terms of the displacement at the cell centre/centroid $\bb{u}_P$ and at the centres of the neighbouring cells $N_i$.

@@

The discretisation is complete but, in its current form, is known to suffer from zero-energy modes, i.e. checkerboarding oscillations.
Here, a Rhie-Chow-type stabilisation term \cite{rhie_numerical_1983} is added to the residual (Equation \ref{eqn:residual}) to quell such oscillations.
The Rhie-Chow stabilisation term, first used for finite volume solid mechanics by  \citet{demirdzic_numerical_1995}, consists of the numerical difference between a diffusion (Laplacian) term calculated using compact and larger computational stencils.
The term introduces numerical diffusion to the discretisation, which reduces at a third-order rate.
% based on the earlier approach of Rhie and Chow \citet{demirdzic_numerical_1995}.
%One issue encountered with the finite volume method is that the discretisation of the governing conservation of momentum equation ( can be unstable and is known to suffer from checker-boarding errors .
%In order to rectify these issues, the Rhie-Chow stabilisation term \cite{rhie_numerical_1983} as introduced into solid mechanics by  is added to the discretised divergence of the stress in equation \ref{eqn:MomentumImplicitExplicit}.
In the current approach, the Rhie-Chow stabilisation term $\mathcal{D}_{\text {Rhie-Chow }}$ for a cell $P$ takes the following form:
\begin{equation} \label{eq:RhieChow}
	\mathcal{D}_{\text {Rhie-Chow}}
	= \sum_{f \in N_f} \alpha \bar{K}_f
	\left[
	\left|\bb{\Delta}_f\right| \frac{ \bb{u}_{N_f} - \bb{u}_P}{\left|\bb{d}_f\right|}
	- \bb{\Delta}_f \cdot \left(\bb{\nabla} \bb{u} \right)_f
	\right]
	\left|\bb{\Gamma}_{f}\right| 
\end{equation}
%which comes from the difference between Equations \ref{eq:diffusion} and \ref{eq:diffusion_exp}.
where $\alpha > 0$ is a user-defined parameter for globally scaling the amount of stabilisation.
Parameter $\bar{K}_f$ is a stiffness-type parameter that gives the stabilisation an appropriate scale and dimension.
Here, $\bar{K}_f = \frac{4}{3}\mu + \kappa = 2\mu + \lambda$ following previous work \cite{Jasak2008, Cardiff, etc}, where $\mu$ is the shear modulus (first Lam\'{e} parameter), $\kappa$ is the bulk modulus, and $\lambda$ is the second Lam\'{e} parameter.
%where $N_f$ represents the set of faces $f$ in cell $P$, and neighbouring cell centre $N_f$ shares face $f$ with the cell $P$.
Vector $\bb{d}_{f}$ connects cell centre $P$ with the other cell sharing face $f$, and $\bb{n}_{f}$ is the outward-facing unit normal to the face $f$.
The vector $\bb{\Delta}_{f} = \frac{\bb{d}_{f}}{\bb{d}_{f} \cdot \bb{n}_{f}}$ is termed the \emph{over-relaxed orthogonal} vector \cite{Jasak1996} and increases in magnitude as the deviation between the $\bb{d}_{f}$ and $\bb{n}_{f}$ vectors increases.
In this way, the amount of stabilisation increases on distorted meshes.
\hl{Should we mention Nishikawa alpha scheme?} \hl{Very similar: but scales differently with mesh distortion}
%and non-orthogonal correction vector $\bb{k}_f=\bb{n}_f-\bb{\Delta}_f$, where $\bb{n}_f$ is the outward-facing unit normal to the face $f$.
%Vector $\bb{d}_f$ connects the centre of cell $P$ with the centre of cell $N_f$ in the updated configuration.
%The first term on the right-hand side is treated implicitly, while the second term - representing non-orthogonal corrections at the face - is treated in a deferred correction manner.

@@

The \emph{inner} linear sparse systems (Equations \ref{eq:segX}, \ref{eq:segY} and \ref{eq:segZ}) can be solved using any typical direct or iterative linear solver approach; however, an incomplete Cholesky pre-conditioned conjugate gradient method \cite{jacobs_generalization_1986} is often preferred as the weakly diagonally dominant characteristic leads to good convergence characteristics.
Algebraic multigrid can be used to accelerate convergence.
%In non-linear problems, this system of equations is solved multiple times with updated coefficients in a fixed-point iteration scheme. 
%As noted in previous articles on segregated methods, the inner system need not be solved to a tight tolerance as coefficients and source terms are approximated from the previous increment; a reduction in the residuals of one order of magnitude is typically sufficient. The outer iterations are performed until the predefined tolerance, typically $1 \times 10^{-6}$, has been achieved \cite{cardiff_lagrangian_2017}. 
%In the current updated Lagrangian approach, the mesh is moved to the deformed configuration at the end of each time step rather than after each outer iteration.
%Since the displacements are calculated at the cell centres, a linear least-squared method is employed here \cite{cardiff_lagrangian_2017} to interpolate the displacement increments to the mesh vertices, allowing the mesh to be moved.
%In this method, a linear least squares plane is fit through a vertex and its immediately adjacent cell centres. For boundary vertices, boundary face-centre values are also included in the fitting.
%
%The procedures have been implemented and publicly shared within the solids4foam toolbox \citep{Cardiff2018, Tukovic2018} of the open-source OpenFOAM software.

@@

%\hl{1}
%Derive Jv for 2x2 system.
The derivation of this approximation can be shown for a $2 \times 2$ system as \cite{Knoll20024}:
\begin{eqnarray}
	\frac{\mathbf{F}(\mathbf{x} + \epsilon \mathbf{v}) - \mathbf{F}(\mathbf{x})}{\epsilon}
	&=&
	\begin{pmatrix}
	\frac{F_1 (x_1 + \epsilon v_1, x_2 + \epsilon v_2) - F_1 (x_1, x_2)}{\epsilon}\\
	\frac{F_2 (x_1 + \epsilon v_1, x_2 + \epsilon v_2) - F_2 (x_1, x_2)}{\epsilon}
	\end{pmatrix} \notag \\
	&\approx&
	\begin{pmatrix}
	\frac{F_1 (x_1,x_2) + \epsilon v_1 \frac{\partial F_1}{\partial u_1} + \epsilon v_2 \frac{\partial F_1}{\partial u_2} - F_1 (x_1, x_2)}{\epsilon}\\
	\frac{F_2 (x_1, x_2) + \epsilon v_1 \frac{\partial F_2}{\partial u_1} + \epsilon v_2 \frac{\partial F_2}{\partial u_2}  - F_2 (x_1, x_2)}{\epsilon}
	\end{pmatrix} \notag \\
	&\approx&
	\begin{pmatrix}
	v_1 \frac{\partial F_1}{\partial u_1} +  v_2 \frac{\partial F_1}{\partial u_2} \\
	v_1 \frac{\partial F_2}{\partial u_1} + v_2 \frac{\partial F_2}{\partial u_2}
	\end{pmatrix} \notag \\
	&\approx&
	\bb{J} \bb{v}
\end{eqnarray}
where a first-order truncated Taylor series expansion about $\bb{u}$ was used to approximate $\bb{F} (\bb{x} + \epsilon \bb{v})$.
%\hl{1b}
%Choosing $\epsilon$ is important.
As noted above, choosing an appropriate value for $\epsilon$ is non-trivial, and care must be taken to balance truncation error (reduced by decreasing $\epsilon$) and round-off error (increased by decreasing $\epsilon$).

@@

%\hl{3}
%- we use approx J to form precon
%- many precon used in lit, we will consider ILU(N) and MG, where MG is expected to be better, but also LU since direct solvers are popular in FE solid mechanics
In the current work, we proposed to use the compact-stencil approximate Jacobian from the segregated algorithm $\tilde{\bb{J}}$ as the preconditioning matrix $\bb{P}$ for the preconditioned Jacobian-free Newton-Krylov method.
\hl{Comment on literature: used before but not yet for solids}.
This preconditioning approach can be considered as a ``physics-based" preconditioner in the classifications of \cite{Knoll}.
The approach is conceptually similar to an approximation of the Jacobian of a higher-order advection scheme by a compact-stencil lower-order upwind scheme.
A benefit of the proposed approach is that existing segregated frameworks can re-use their existing discretisation and storage implementations.
Concretely, the Jacobian-free Newton-Krylov method requires only a procedure for forming this preconditioning matrix and a procedure for explicitly evaluating the residual.
Both routines are easily implemented in an existing segregated framework.
The only additional required procedure is an interface to an existing Jacobian-free Newton-Krylov implementation.
%"The motivation behind this approach is that there exist numerous, legacy algorithms to solve nonlinear systems, both IVPs and BVPs. These algorithms typically were developed with some insight into the time scales or physical behavior of the problem. As a benefit of this insight, a reduced implicit system, or a sequence of segregated explicit or implicit systems may be solved in place of the fully coupled system. "
In the current work, the PETSc toolbox \cite{PETSc} is used as the nonlinear solver, where driven by a finite volume solver in the OpenFOAM toolbox \cite{OF}.

@@

%\hl{4 - globalisation}
A challenge with Newton-type methods, including Jacobian-free versions, is convergence can be poor when far from the true solution, and divergence is often a real possibility.
Globalisation refers to steering an initial solution towards the quadratic convergence range of the Newton method.
Several strategies are possible, and it is common to combine approaches \cite{Knoell2004}.
In the current work, a line search procedure is used to select the $s$ parameter in the solution update step (the second line in Equations \ref{eq:NewtonRaphson}).
Line search methods assume the Newton update direction is correction and aim to find a scalar $s > 0$ that decreases the residual $\bb{R}(\bb{u}_k + s \delta \bb{u}) < \bb{R}(\bb{u}_k)$.
The scalar $s$ is typically $\leq 1$, but extrapolation ($>1$) is also possible for accelerating convergence, albeit at the expense of robustness.

@@

%\hl{5 - oversolving}
%Comment on over-solving => also applies to the segregated system.
%This could be a parameter we look at.
A final comment on the Jacobian-free Newton-Krylov solution algorithm is the potential importance of \emph{oversolving}.
Here, oversolving refers to solving the linear system to too tight a tolerance during the early Newton iterations, essentially wasting time when the solution is far from the true solution.
In addition, some authors \cite{See164_and_176_in_Knoll2004} have shown Newton convergence to be worse when the earlier iterations are solved to too tight a tolerance.
The concept of oversolving also applies to segregated solution procedures and has been well-known since the early work of Demird\v{z}i\'{c} and co-workers \cite{Demirdzic}, where the residuals are typically reduced by one order of magnitude in the inner linear system.
The optimal choice of residual reduction for a Jacobian-free Newton-Krylov finite volume solid mechanics procedure is explored in Section \ref{sec:test_cases}. \hl{Check: do we examine this?}

@@

The manufactured displacement solution is applied at the domain's boundaries, and inertial effects are neglected.
Two forms of mesh are examined: (i) structured, uniform hexahedra and (ii) regular polyhedra (Figure \ref{fig:mms_mesh}).
To create the polyhedral meshes, structured tetrahedra meshes are first generated using Gmsh \cite{geuzaine2009gmsh} and then converted to their dual polyhedral representations using the OpenFOAM \texttt{polyDualMesh} utility.
Starting from an initial mesh spacing of $0.04$ m, six meshes are created by successively halving the spacing.
The cell numbers (hexahedra and polyhedral meshes) are 125, $1\,000$, $8\,000$, $64\,000$, $512\,000$, and $4\,096\,000$.

@@

The solution domain is taken as one-eighth of a $1 \times 1 \times 1$ m cube centred on the sphere.
The analytical tractions are applied at the far boundaries of the domain to mitigate the effects of finite geometry.
Unstructured polyhedral meshes of varying densities are employed (\hl{state average cell widths, and number of cells in each mesh}).
The mesh with 2,257 cells is shown in Figure \ref{fig:spherical_cavity}.
Initially, unstructured tetrahedral meshes were generated using the Gmsh meshing utility \cite{geuzaine2009gmsh}, followed by conversion to their dual polyhedral representations using the OpenFOAM \texttt{polyDualMesh} utility.

\paragraph{Case 2: Out-of-plane bending of an elliptic plate}
This 3-D, static, linear elastic test case (Figure \ref{fig:elliptic_plate}) consists of a thick elliptic plate (0.6 m thick) with a centred elliptic hole, with the inner and outer ellipses given as
\begin{eqnarray}
	\left(\frac{x}{2}\right)^2 + \left(\frac{y}{1}\right)^2 = 1 & \text{inner ellipse} \\
	\left(\frac{x}{3.25}\right)^2 + \left(\frac{y}{2.75}\right)^2 = 1 & \text{outer ellipse}
\end{eqnarray}
The case has been described by the National Agency for Finite Element Methods and Standards (NAFEMS) \cite{NAFEMS}, and analysed using finite volume procedures by \citet{Demirdzic} and \citet{Cardiff2016}.
Symmetry allows one quarter of the geometry to be simulated.
A constant pressure of 1 MPa is applied to the upper surface, and the outer surface is fully clamped.
The mechanical properties are: $E = 210$ GPa, $\nu = 0.3$.
\hl{Comment on the meshes used}.
\begin{figure}[htbp]
   \centering
   %\includegraphics[width=0.2\textwidth]{figures/elliptic_plate.pdf} 
   \includegraphics[width=0.2\textwidth]{figures/placeholder.pdf} 
   \caption{Elliptic Plate case geometry, mesh and loading conditions}
   \label{fig:elliptic_plate}
\end{figure}

>>>>

Finite volume formulations for solid mechanics are heavily influenced by their fluid mechanics counterparts, favouring segregated implicit and fully explicit methods.
Segregated approaches, where the governing momentum equation is temporarily decomposed into scalar component equations, offer memory efficiency and simplicity of implementation, but the outer coupling Picard iterations often suffer from slow convergence.
Explicit formulations are straightforward to implement and offer superior robustness but are only efficient for high-speed dynamics, where the physics requires small time increments.
In contrast, the finite element community commonly employs Newton-Raphson-type solution algorithms, which necessitate repeated assembly of the Jacobian matrix and solution of the resulting block-coupled non-diagonally dominant linear system.
A disadvantage of traditional Newton-based approaches is that they typically require explicit formation and storage of the Jacobian matrix, which can be computationally expensive and memory-intensive.
A further disadvantage from a finite volume perspective is that extending existing code frameworks from segregated algorithms to a coupled Newton-Raphson-type approach is challenging in terms of the required assembly, storage, and solution of the resulting block-coupled system.
In addition, the derivation of the true Jacobian matrix is non-trivial.
Consequently, similar block-coupled solution finite volume methods are rare in the literature \citep{Das2011, Cardiff2016, Castrillo2024}.
The motivation of the current work is to seek (or exceed) the robustness and efficiency of block-coupled Newton-Raphons approaches in a way that can be easily incorporated into existing segregated solution frameworks.
To this end, the current article examines the efficacy of \emph{Jacobian-free} Newton-Krylov methods, where the quadratic convergence of Newton methods can potentially be achieved without deriving, assembling and storing the exact Jacobian.

@@

A crucial aspect of ensuring the efficiency and robustness of the Jacobian-free Newton-Krylov method is the choice of a suitable preconditioner for the Krylov iterations.
This preconditioner is often derived from the exact Jacobian matrix in traditional Newton methods.
However, the Jacobian-free approach does not allow direct access to the full Jacobian matrix, necessitating an alternative strategy to approximate its action.
To this end, and to extend existing segregated frameworks, we propose using a compact-stencil approximate Jacobian as the preconditioner. This approximate Jacobian corresponds to the matrix typically employed in segregated approaches; similar approaches are successful in fluid mechanics applications \citep{Nishikawa2020, nonNewtonianJFNKPaper}; however, it is unclear if such an approach is suitable for solid mechanics - a question which we hope to answer in this work.
By leveraging this compact-stencil approximate Jacobian, we aim to effectively precondition the Krylov iterations, enhancing convergence while maintaining the memory and computational savings that define the Jacobian-free and segregated methods.
Similarly, if such an approach is efficient, it would naturally fit into existing segregated frameworks, as existing matrix storage and assembly can be reused.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Cell-Centred Finite Volume Discretisation}
\label{sec:discretisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this work, a nominally second-order cell-centred finite volume discretisation is employed, as described previously, for example, \citep{Cardiff2017, batistic2022, tukovic2013, jasak2000}.
Consequently, only a summary of the discretisation is presented below.

@@

The conservation equation (Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL}, or \ref{eqn:momentum_UL}) is applied to each cell (control volume) in the computational mesh and discretised in terms of the displacement at the cell centre/centroid $\bb{u}_P$ and at the centres of the neighbouring cells $N_i$.

@@

The discretisation is complete but, in its current form, is known to suffer from zero-energy modes, i.e. checkerboarding oscillations.
Here, a Rhie-Chow-type stabilisation term \cite{Rhie1983} is added to the residual (Equation \ref{eqn:residual}) to quell such oscillations.
The Rhie-Chow stabilisation term, first used for finite volume solid mechanics by  \citet{Demirdzic1995}, consists of the numerical difference between a diffusion (Laplacian) term calculated using compact and larger computational stencils.
The term introduces numerical diffusion to the discretisation, which reduces at a third-order rate.
% based on the earlier approach of Rhie and Chow \citet{demirdzic_numerical_1995}.
%One issue encountered with the finite volume method is that the discretisation of the governing conservation of momentum equation ( can be unstable and is known to suffer from checker-boarding errors .
%In order to rectify these issues, the Rhie-Chow stabilisation term \cite{rhie_numerical_1983} as introduced into solid mechanics by  is added to the discretised divergence of the stress in equation \ref{eqn:MomentumImplicitExplicit}.
In the current approach, the Rhie-Chow stabilisation term $\mathcal{D}_{\text {Rhie-Chow }}$ for a cell $P$ takes the following form:
\begin{equation} \label{eq:RhieChow}
	\mathcal{D}_{\text {Rhie-Chow}}
	= \sum_{f \in N_f} \alpha \bar{K}_f
	\left[
	\left|\bb{\Delta}_f\right| \frac{ \bb{u}_{N_f} - \bb{u}_P}{\left|\bb{d}_f\right|}
	- \bb{\Delta}_f \cdot \left(\bb{\nabla} \bb{u} \right)_f
	\right]
	\left|\bb{\Gamma}_{f}\right| 
\end{equation}
%which comes from the difference between Equations \ref{eq:diffusion} and \ref{eq:diffusion_exp}.
where $\alpha > 0$ is a user-defined parameter for globally scaling the amount of stabilisation.
Parameter $\bar{K}_f$ is a stiffness-type parameter that gives the stabilisation an appropriate scale and dimension.
Here, $\bar{K}_f = \frac{4}{3}\mu + \kappa = 2\mu + \lambda$ following previous work \cite{Jasak2008, Cardiff, etc}, where $\mu$ is the shear modulus (first Lam\'{e} parameter), $\kappa$ is the bulk modulus, and $\lambda$ is the second Lam\'{e} parameter.
%where $N_f$ represents the set of faces $f$ in cell $P$, and neighbouring cell centre $N_f$ shares face $f$ with the cell $P$.
Vector $\bb{d}_{f}$ connects cell centre $P$ with the other cell sharing face $f$, and $\bb{n}_{f}$ is the outward-facing unit normal to the face $f$.
The vector $\bb{\Delta}_{f} = \frac{\bb{d}_{f}}{\bb{d}_{f} \cdot \bb{n}_{f}}$ is termed the \emph{over-relaxed orthogonal} vector \cite{Jasak1996} and increases in magnitude as the deviation between the $\bb{d}_{f}$ and $\bb{n}_{f}$ vectors increases.
In this way, the amount of stabilisation increases on distorted meshes.
\hl{Should we mention Nishikawa alpha scheme?} \hl{Very similar: but scales differently with mesh distortion}
%and non-orthogonal correction vector $\bb{k}_f=\bb{n}_f-\bb{\Delta}_f$, where $\bb{n}_f$ is the outward-facing unit normal to the face $f$.
%Vector $\bb{d}_f$ connects the centre of cell $P$ with the centre of cell $N_f$ in the updated configuration.
%The first term on the right-hand side is treated implicitly, while the second term - representing non-orthogonal corrections at the face - is treated in a deferred correction manner.

@@

The \emph{inner} linear sparse systems (Equations \ref{eq:segX}, \ref{eq:segY} and \ref{eq:segZ}) can be solved using any typical direct or iterative linear solver approach; however, an incomplete Cholesky pre-conditioned conjugate gradient method \cite{Jacobs1986} is often preferred as the weakly diagonally dominant characteristic leads to good convergence characteristics.
Algebraic multigrid can be used to accelerate convergence.
%In non-linear problems, this system of equations is solved multiple times with updated coefficients in a fixed-point iteration scheme. 
%As noted in previous articles on segregated methods, the inner system need not be solved to a tight tolerance as coefficients and source terms are approximated from the previous increment; a reduction in the residuals of one order of magnitude is typically sufficient. The outer iterations are performed until the predefined tolerance, typically $1 \times 10^{-6}$, has been achieved \cite{cardiff_lagrangian_2017}. 
%In the current updated Lagrangian approach, the mesh is moved to the deformed configuration at the end of each time step rather than after each outer iteration.
%Since the displacements are calculated at the cell centres, a linear least-squared method is employed here \cite{cardiff_lagrangian_2017} to interpolate the displacement increments to the mesh vertices, allowing the mesh to be moved.
%In this method, a linear least squares plane is fit through a vertex and its immediately adjacent cell centres. For boundary vertices, boundary face-centre values are also included in the fitting.
%
%The procedures have been implemented and publicly shared within the solids4foam toolbox \citep{Cardiff2018, Tukovic2018} of the open-source OpenFOAM software.

@@

%\hl{1}
%Derive Jv for 2x2 system.
The derivation of this approximation can be shown for a $2 \times 2$ system as \cite{Knoll2004}:
\begin{eqnarray}
	\frac{\mathbf{F}(\mathbf{x} + \epsilon \mathbf{v}) - \mathbf{F}(\mathbf{x})}{\epsilon}
	&=&
	\begin{pmatrix}
	\frac{F_1 (x_1 + \epsilon v_1, x_2 + \epsilon v_2) - F_1 (x_1, x_2)}{\epsilon}\\
	\frac{F_2 (x_1 + \epsilon v_1, x_2 + \epsilon v_2) - F_2 (x_1, x_2)}{\epsilon}
	\end{pmatrix} \notag \\
	&\approx&
	\begin{pmatrix}
	\frac{F_1 (x_1,x_2) + \epsilon v_1 \frac{\partial F_1}{\partial u_1} + \epsilon v_2 \frac{\partial F_1}{\partial u_2} - F_1 (x_1, x_2)}{\epsilon}\\
	\frac{F_2 (x_1, x_2) + \epsilon v_1 \frac{\partial F_2}{\partial u_1} + \epsilon v_2 \frac{\partial F_2}{\partial u_2}  - F_2 (x_1, x_2)}{\epsilon}
	\end{pmatrix} \notag \\
	&\approx&
	\begin{pmatrix}
	v_1 \frac{\partial F_1}{\partial u_1} +  v_2 \frac{\partial F_1}{\partial u_2} \\
	v_1 \frac{\partial F_2}{\partial u_1} + v_2 \frac{\partial F_2}{\partial u_2}
	\end{pmatrix} \notag \\
	&\approx&
	\bb{J} \bb{v}
\end{eqnarray}
where a first-order truncated Taylor series expansion about $\bb{u}$ was used to approximate $\bb{F} (\bb{x} + \epsilon \bb{v})$.
%\hl{1b}
%Choosing $\epsilon$ is important.
As noted above, choosing an appropriate value for $\epsilon$ is non-trivial, and care must be taken to balance truncation error (reduced by decreasing $\epsilon$) and round-off error (increased by decreasing $\epsilon$).

@@

%\hl{3}
%- we use approx J to form precon
%- many precon used in lit, we will consider ILU(N) and MG, where MG is expected to be better, but also LU since direct solvers are popular in FE solid mechanics
In the current work, we proposed to use the compact-stencil approximate Jacobian from the segregated algorithm $\tilde{\bb{J}}$ as the preconditioning matrix $\bb{P}$ for the preconditioned Jacobian-free Newton-Krylov method.
\hl{Comment on literature: used before but not yet for solids}.
This preconditioning approach can be considered as a ``physics-based" preconditioner in the classifications of \cite{Knoll2004}.
The approach is conceptually similar to an approximation of the Jacobian of a higher-order advection scheme by a compact-stencil lower-order upwind scheme.
A benefit of the proposed approach is that existing segregated frameworks can re-use their existing discretisation and storage implementations.
Concretely, the Jacobian-free Newton-Krylov method requires only a procedure for forming this preconditioning matrix and a procedure for explicitly evaluating the residual.
Both routines are easily implemented in an existing segregated framework.
The only additional required procedure is an interface to an existing Jacobian-free Newton-Krylov implementation.
%"The motivation behind this approach is that there exist numerous, legacy algorithms to solve nonlinear systems, both IVPs and BVPs. These algorithms typically were developed with some insight into the time scales or physical behavior of the problem. As a benefit of this insight, a reduced implicit system, or a sequence of segregated explicit or implicit systems may be solved in place of the fully coupled system. "
In the current work, the PETSc toolbox \cite{PETSc} is used as the nonlinear solver, where driven by a finite volume solver in the OpenFOAM toolbox \cite{OF}.

@@

%\hl{4 - globalisation}
A challenge with Newton-type methods, including Jacobian-free versions, is convergence can be poor when far from the true solution, and divergence is often a real possibility.
Globalisation refers to steering an initial solution towards the quadratic convergence range of the Newton method.
Several strategies are possible, and it is common to combine approaches \cite{Knoll2004}.
In the current work, a line search procedure is used to select the $s$ parameter in the solution update step (the second line in Equations \ref{eq:NewtonRaphson}).
Line search methods assume the Newton update direction is correction and aim to find a scalar $s > 0$ that decreases the residual $\bb{R}(\bb{u}_k + s \delta \bb{u}) < \bb{R}(\bb{u}_k)$.
The scalar $s$ is typically $\leq 1$, but extrapolation ($>1$) is also possible for accelerating convergence, albeit at the expense of robustness.

@@

%\hl{5 - oversolving}
%Comment on over-solving => also applies to the segregated system.
%This could be a parameter we look at.
A final comment on the Jacobian-free Newton-Krylov solution algorithm is the potential importance of \emph{oversolving}.
Here, oversolving refers to solving the linear system to too tight a tolerance during the early Newton iterations, essentially wasting time when the solution is far from the true solution.
In addition, some authors %\cite{See164_and_176_in_Knoll2004}
\cite{Knoll2004} have shown Newton convergence to be worse when the earlier iterations are solved to too tight a tolerance.
The concept of oversolving also applies to segregated solution procedures and has been well-known since the early work of Demird\v{z}i\'{c} and co-workers \cite{Demirdzic}, where the residuals are typically reduced by one order of magnitude in the inner linear system.
The optimal choice of residual reduction for a Jacobian-free Newton-Krylov finite volume solid mechanics procedure is explored in Section \ref{sec:test_cases}. \hl{Check: do we examine this?}

@@

The manufactured displacement solution is applied at the domain's boundaries, and inertial effects are neglected.
Two forms of mesh are examined: (i) structured, uniform hexahedra and (ii) regular polyhedra (Figure \ref{fig:mms_mesh}).
To create the polyhedral meshes, structured tetrahedra meshes are first generated using Gmsh \cite{Geuzaine2009} and then converted to their dual polyhedral representations using the OpenFOAM \texttt{polyDualMesh} utility.
Starting from an initial mesh spacing of $0.04$ m, six meshes are created by successively halving the spacing.
The cell numbers (hexahedra and polyhedral meshes) are 125, $1\,000$, $8\,000$, $64\,000$, $512\,000$, and $4\,096\,000$.

@@

The solution domain is taken as one-eighth of a $1 \times 1 \times 1$ m cube centred on the sphere.
The analytical tractions are applied at the far boundaries of the domain to mitigate the effects of finite geometry.
Unstructured polyhedral meshes of varying densities are employed (\hl{state average cell widths, and number of cells in each mesh}).
The mesh with 2,257 cells is shown in Figure \ref{fig:spherical_cavity}.
Initially, unstructured tetrahedral meshes were generated using the Gmsh meshing utility \cite{Geuzaine2009}, followed by conversion to their dual polyhedral representations using the OpenFOAM \texttt{polyDualMesh} utility.

\paragraph{Case 2: Out-of-plane bending of an elliptic plate}
This 3-D, static, linear elastic test case (Figure \ref{fig:elliptic_plate}) consists of a thick elliptic plate (0.6 m thick) with a centred elliptic hole, with the inner and outer ellipses given as
\begin{eqnarray}
	\left(\frac{x}{2}\right)^2 + \left(\frac{y}{1}\right)^2 = 1 & \text{inner ellipse} \\
	\left(\frac{x}{3.25}\right)^2 + \left(\frac{y}{2.75}\right)^2 = 1 & \text{outer ellipse}
\end{eqnarray}
The case has been described by the National Agency for Finite Element Methods and Standards (NAFEMS) \cite{NAFEMS}, and analysed using finite volume procedures by \citet{Demirdzic1997} and \citet{Cardiff2016}.
Symmetry allows one quarter of the geometry to be simulated.
A constant pressure of 1 MPa is applied to the upper surface, and the outer surface is fully clamped.
The mechanical properties are: $E = 210$ GPa, $\nu = 0.3$.
\hl{Comment on the meshes used}.
\begin{figure}[htbp]
   \centering
   %\includegraphics[width=0.2\textwidth]{figures/elliptic_plate.pdf} 
   \includegraphics[width=0.2\textwidth]{figures/placeholder.pdf} 
   \caption{Elliptic Plate case geometry, mesh and loading conditions}
   \label{fig:elliptic_plate}
\end{figure}
